Iconic representations of molluscs through the ages

### 2012-10-25 large_number_squeezing_with_atoms.tex
Add more detail to two-mode section

@@

@@

This system can be solved analytically, and expressions for quadrature, absolute and relative number squeezing can be derived. We present the solutions here; the full derivation can be found in \ref{appendixTwoModeDerivation}.
If we define
\begin{eqnarray}
s &=& \sin(\Omega (t_3 - t_2)) \\
c &=& \cos(\Omega (t_3 - t_2)) \\
\lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}} / \hbar \\
A &=& \sqrt{n_a} \exp [n_a (e^{-i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
A_2 &=& n_a \exp [n_a (e^{-2i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
B &=& -i e^{i \phi}\sqrt{n_b} \exp [n_b (e^{-i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
D &=& AB^* - A^*B
\end{eqnarray}
then the expectation values for number, number variance, and number difference variance after appling the final coupling field for a time $\tau = t_3-t_2$ are given by
\begin{eqnarray}
N_a(t_3) &=& n_a c^2 + n_b s^2 + i c s D \\
%
N_b(t_3) &=& n_a s^2 + n_b c^2 - i c s D \\
%
{\mathrm{Var}} [ N_a](t_3) &=& n_a c^2 + n_b s^2 + 2 n_a n_b c^2 s^2 \\
       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2) \nonumber \\
       && + i c s D (1-2 n_a c^2 -2 n_b s^2)  \nonumber\\
       && + 2 i c^3 s n_a (e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - e^{i(\lambda_{aa} - \lambda_{ab})} A^* B ) \nonumber \\
       && + 2 i c s^3 n_b (e^{i(\lambda_{bb} - \lambda_{ab})} A B^* - e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B ) \nonumber\\
%
{\mathrm{Var}} [ N_b](t_3) &=&  n_a s^2 + n_b c^2 + 2 n_a n_b c^2 s^2 \nonumber \\
       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2 ) \nonumber \\
       && + i c s D (-1+2 n_a s^2 + 2 n_b c^2)  \nonumber \\
       && + 2 i c^3 s n_b (e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - e^{i(\lambda_{bb} - \lambda_{ab})} A B^* ) \nonumber \\
       && + 2 i c s^3 n_a (e^{i(\lambda_{aa} - \lambda_{ab})} A^* B - e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* ) \\
%
{\mathrm{Var}} [ N_a - N_b](t_3) &=& n_a + n_b + 4 i c s D (n_a - n_b)(s^2 - c^2) \nonumber \\
       && + 4 c^2 s^2 (2 n_a n_b +D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} A_2^* B_2 - e^{-i(\lambda_{aa} - \lambda_{bb})} A_2 B_2^* ) \nonumber \\
       && + 4 i c s (c^2 - s^2) (n_a e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - n_a e^{i(\lambda_{aa} - \lambda_{ab})} A^* B 
                     + n_b e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - n_b e^{i(\lambda_{bb} - \lambda_{ab})} A B^*)
\end{eqnarray}

@@

\section{Meanfield multimode problems}
\begin{itemize}
  \item Point out that different densities have phases that evolve at different rates
  \item This leads to poor mode matching when recombining if two states have different nonlinearity*number values 
  \item give a plot of the mode in |a> and |b> at the beginning and end of the hold time to show mismatch
  \item This means constant density is good since mode shape doesn't change
  \item Plot of squeezing for given set of params for gaussian, TF, constant density, show it get steadily better
  \item Say we can use a pi pulse to swap the populations half way through, so if the states have different nonlinearities we still end up with the same mode shapes when we recombine. The fact that this works will be clear from the SM equations for a(t), b(t) when the beam splitter is on
  \item show some plots of squeezing with and without the pi pulse
\end{itemize}

>>>>

{\bf{Maybe expand on the link between a full BEC in a trap and a the following hamiltonian?}}

@@

In another interpretation, the sequence of pulses results in a Ramsey interferometer, with a final beam splitter phase of $\phi$.

@@

This system can be solved analytically, and expressions for quadrature, absolute and relative number squeezing can be derived. We present the solutions here; the full derivation can be found in \ref{appendixTwoModeDerivation}.
If we define
\begin{eqnarray}
s &=& \sin(\Omega (t_3 - t_2)) \\
c &=& \cos(\Omega (t_3 - t_2)) \\
\lambda_{ij} &=& \chi_{ij} \tau_{\mathrm{hold}} / \hbar \\
A &=& \sqrt{n_a} \exp [n_a (e^{-i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
A_2 &=& n_a \exp [n_a (e^{-2i(\lambda_{aa}-\lambda_{ab})} - 1 )] \\
B &=& -i e^{i \phi}\sqrt{n_b} \exp [n_b (e^{-i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
B_2 &=& -n_b e^{2i\phi} \exp [n_b (e^{-2i(\lambda_{bb}-\lambda_{ab})} - 1 )] \\
D &=& AB^* - A^*B
\end{eqnarray}
then the expectation values for number, number variance, and number difference variance after appling the final coupling field for a time $\tau = t_3-t_2$ are given by
\begin{eqnarray}
N_a(t_3) &=& n_a c^2 + n_b s^2 + i c s D \\
%
N_b(t_3) &=& n_a s^2 + n_b c^2 - i c s D \\
%
{\mathrm{Var}} [ N_a](t_3) &=& n_a c^2 + n_b s^2 + 2 n_a n_b c^2 s^2 \\
       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2) \nonumber \\
       && + i c s D (1-2 n_a c^2 -2 n_b s^2)  \nonumber\\
       && + 2 i c^3 s n_a (e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - e^{i(\lambda_{aa} - \lambda_{ab})} A^* B ) \nonumber \\
       && + 2 i c s^3 n_b (e^{i(\lambda_{bb} - \lambda_{ab})} A B^* - e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B ) \nonumber\\
%
{\mathrm{Var}} [ N_b](t_3) &=&  n_a s^2 + n_b c^2 + 2 n_a n_b c^2 s^2 \nonumber \\
       && + c^2 s^2 (D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} B_2 A_2^* - e^{-i(\lambda_{aa} - \lambda_{bb})} B_2^* A_2 ) \nonumber \\
       && + i c s D (-1+2 n_a s^2 + 2 n_b c^2)  \nonumber \\
       && + 2 i c^3 s n_b (e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - e^{i(\lambda_{bb} - \lambda_{ab})} A B^* ) \nonumber \\
       && + 2 i c s^3 n_a (e^{i(\lambda_{aa} - \lambda_{ab})} A^* B - e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* ) \\
%
{\mathrm{Var}} [ N_a - N_b](t_3) &=& n_a + n_b + 4 i c s D (n_a - n_b)(s^2 - c^2) \nonumber \\
       && + 4 c^2 s^2 (2 n_a n_b +D^2 - e^{i(\lambda_{aa} - \lambda_{bb})} A_2^* B_2 - e^{-i(\lambda_{aa} - \lambda_{bb})} A_2 B_2^* ) \nonumber \\
       && + 4 i c s (c^2 - s^2) (n_a e^{-i(\lambda_{aa} - \lambda_{ab})} A B^* - n_a e^{i(\lambda_{aa} - \lambda_{ab})} A^* B 
                     + n_b e^{-i(\lambda_{bb} - \lambda_{ab})} A^* B - n_b e^{i(\lambda_{bb} - \lambda_{ab})} A B^*)
\end{eqnarray}
In the case where all the nonlinearities are equal, such that $\chi_{aa} = \chi_{ab} = \chi_{bb}$, the number variances all reduce to those of a coherent state, e.g. ${\mathrm{Var}}[N_a]=N_a$, an no number squeezing is possible.

There are three independent nonlinearities $\chi_{ij}$ in the Hamiltonian (\ref{eqTwoModeHamiltonian}). In the expressions for the number variances, however, the nonlinearities only appear in the cyclic combinations $\chi_{aa}-\chi_{ab}$, $\chi_{bb}-\chi_{ab}$ and $\chi_{aa}-\chi_{bb}$, meaning that number squeezing is parameterized by only two independent quantities which will be linear combinations of the nonlinearities. For example, since $\chi_{bb}-\chi_{ab} = (\chi_{aa}-\chi_{ab}) - (\chi_{aa}-\chi_{bb})$ it is possible to parameterize the number squeezing entirely in terms of the nonlinear differences $\chi_{aa}-\chi_{ab}$ and $\chi_{aa}-\chi_{bb}$.

{\bf{There seems to be this bogus rule of thumb in the literature that the amount of squeezing you can get is related to $\chi_{11}+\chi_{22}-2\chi_{12}$. Is it worth pointing out that this is wrong?}}

While the analytic expressions for the number variances are not particularly transparent, it is easy to evaluate and plot them for any given set of parameters. As an example, in Figure \ref{figxxx} we plot the solution for the normalized number variance in mode $|a\rangle$ for a particular set of nonlinearities, choosing a number of different values of $\phi$. The plots illustrate the basic features of the problem: During the final coupling period, the number variance undergoes cyclic variation, with a series of minima. It is necessary to choose both the correct phase $\phi$ as well as the correct length of time for the coupling to be applied in order to obtain optimum squeezing.

%Can Uaa + Ubb - 2Uab be zero and still get squeezing? Yep, Uaa=0.03, Ubb=0.01, Uab=0.2 gives squeezing for all quantities of interest

@@

\section{Meanfield multimode problems}
\begin{itemize}
  \item Point out that different densities have phases that evolve at different rates
  \item This leads to poor mode matching when recombining if two states have different nonlinearity*number values 
  \item give a plot of the mode in $|a\rangle$ and $|b\rangle$ at the beginning and end of the hold time to show mismatch
  \item This means constant density is good since mode shape doesn't change
  \item Plot of squeezing for given set of params for gaussian, TF, constant density, show it get steadily better
  \item Say we can use a pi pulse to swap the populations half way through, so if the states have different nonlinearities we still end up with the same mode shapes when we recombine. The fact that this works will be clear from the SM equations for a(t), b(t) when the beam splitter is on
  \item show some plots of squeezing with and without the pi pulse
\end{itemize}
